### 2024-12-18 analysis_objects/README.Rmd

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Analysis objects

This folder contains all [analysis objects](https://github.com/StefanVermeent/liss_ef_2024/tree/main/analysis_objects), i.e., the outputs of the various data processing scripts.
The analysis objects are loaded by the `manuscript.qmd` and `supplement.qmd` file to fill in the statistics, Tables, and Figures in the manuscript and supplemental materials