### 2024-12-18 materials/README.Rmd

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

## Materials {#materials}

This folder contains all the materials of the cognitive tasks.
All tasks were programmed in [JsPsych 7.3.2](https://www.jspsych.org/7.3/).

1.  [`pilot01/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/pilot01): Folder containing the pilot materials for the Flanker, Simon, Posner, Animacy-size, Color-shape, and Global-local tasks. Note that the Global-local task used in this pilot study was different from the one used in the manuscript.
2.  [`pilot02/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/pilot02): Folder containing the pilot materials for the alternative Global-local task (as used in the manuscript).
3.  [`pilot03/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/pilot03): Folder containing the pilot materials for all six tasks used in the manuscript.
4.  [`liss/`](https://github.com/StefanVermeent/liss_wm_profiles_2023/tree/master/materials/1_liss): Folder containing the final materials of all six tasks as used in the main study.

Within each study folder, each task also has it's own folder.
In each folder, you will find two files:
`0_[task]_setup.js` contains the Javascript code for functions and stimulus set-up.
`1_[task]_instructions.js` contains the Javascript code for all task instructions.
`2_[task]_procedure.js` contains Javascript code defining the stimuli and trials for both practice and test blocks.

In each of these study folders, you will find several additional files and folders:

1.  `index.html`. Script which combines the trial blocks and initiates the tasks.
2.  `img/`. Folder containing stimulus images.
4.  `scripts`. Scripts handling trial rules.

**Run tasks locally.** To run the tasks locally on your device, navigate to any task folder and open the `index.html` file in a browser. 
This will automatically start the cognitive task.

**Run tasks online.** The tasks are hosted online at [https://liss-ef-2024-pilot3.netlify.app/](https://liss-ef-2024-pilot3.netlify.app/).
Navigating to this url will automatically start the tasks.