### 2024-12-18 scripts/README.Rmd
add script README

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Data Processing Scripts

There are four types of R-scripts in this repository, each with a separate folder.

-   [`0_functions/`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/scripts/0_functions): Custom R-functions written for this project.
-   [`0_pilot/`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/scripts/0_pilot): Scripts analyzing the Pilot data.
-   [`1_study/`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/scripts/1_study): The full pipeline for cleaning data, running models, and conducting all analyses.

Each script takes an input(s) and produces output(s).
All outputs are stored in [analysis_objects/](https://github.com/StefanVermeent/liss_ef_2024/tree/main/analysis_objects).
These outputs include list with exclusions, DDM models, and SEM results.

The analysis objects, in turn, are used to stage all the statistics, Tables, and Figures included in the manuscript and supplemental material.
This is done in [`manuscript/scripts/staging.R`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/manuscript/scripts/staging.R) and in [`supplement/stage_sup.R`](https://github.com/StefanVermeent/liss_ef_2024/tree/main/supplement/stage_sup.R).